---
title: "Appendix"
author: "Jiawei Meng"
output: pdf_document
---
__Question 1: Diarrheal Prevalence__
```{r}
### import data
diarrhea <- read.csv ("/Users/mengjiawei/Desktop/materials/2020fall/525/Final/diarrhea.csv",header=T)

# check observations
dim(diarrhea)

# check missing data
summary(diarrhea)

# the number of school in the data
length(unique(diarrhea$School))

# the number of observations in each school
table(diarrhea$School)

### descriptive statistics
diarrhea$Sex<-factor(diarrhea$Sex,levels=c(0,1),
                    labels = c("boy","girl"))
diarrhea$treat<-factor(diarrhea$treat,levels=c(0,1),
                    labels = c("no","yes"))
diarrhea$Z<-factor(diarrhea$Z,levels=c(0,1),
                    labels = c("no","yes"))
diarrhea$Sex
library(table1)
label(diarrhea$Sex)<-"Student’s sex"
label(diarrhea$School)<-"School ID"
label(diarrhea$Age)<-"Student’s age at the time of survey"
label(diarrhea$treat)<-"the presence of a school-level program to improve sanitation and hygiene"
label(diarrhea$Z)<-"whether the child had diarrhea over the last three days"
table1(~ Sex + Age + Z |treat, data = diarrhea)

diarrhea$treat
library(tableone)
listvars = c("Sex", "Age","Z")

catvars = c("Sex","Z")
diarrhea$treat
table1 = CreateTableOne(vars=listvars, data=diarrhea, factorVars = catvars, includeNA = T, strata = "treat")
tbl=print(table1, nonnormal=c("Sex","Z"))
knitr::kable(tbl)
install.packages("compareGroups")
library(compareGroups)
restab <- createTable(tbl, digits = c(p14 = 0, hormo=1), type = 1, 
                      hide.no = "no", show.n = TRUE)
```

```{R}
# instal required package
library(gee)
library(geepack)
# load data and factor variable
diarrhea <- read.csv ("/Users/mengjiawei/Desktop/2020fall/525/Final/diarrhea.csv",header=T)

diarrhea$Sex<-factor(diarrhea$Sex,levels=c(0,1),
                    labels = c("boy","girl"))
diarrhea$treat<-factor(diarrhea$treat,levels=c(0,1),
                    labels = c("no","yes"))
# check interaction
#treat
fittrt = gee(Z~treat ,family = binomial(link = "logit"),data=diarrhea, 
                id=School, corstr = "exchangeable")
2* pnorm(abs(coef(summary(fittrt))[,5]), lower.tail = FALSE)

#age
fitage = gee(Z~treat + Age,family = binomial(link = "logit"),data=diarrhea, 
                id=School, corstr = "exchangeable")
2* pnorm(abs(coef(summary(fitage))[,5]), lower.tail = FALSE)

fitagei = gee(Z~treat * Age,family = binomial(link = "logit"),data=diarrhea, 
                 id=School, corstr = "exchangeable")
2* pnorm(abs(coef(summary(fitagei))[,5]), lower.tail = FALSE)

#sex
fitsex = gee(Z~ treat + Sex,family = binomial(link = "logit"),data=diarrhea, 
                id=School, corstr = "exchangeable")
2* pnorm(abs(coef(summary(fitsex))[,5]), lower.tail = FALSE)
fitsexi = gee(Z~treat * Sex,family = binomial(link = "logit"),data=diarrhea, 
                 id=School, corstr = "exchangeable")
2* pnorm(abs(coef(summary(fitsexi))[,5]), lower.tail = FALSE)

#two inter
fittwo = gee(Z~ treat * Sex + treat* Age,family = binomial(link = "logit"),
                data=diarrhea, id=School, corstr = "exchangeable")
summary(fittwo)
2* pnorm(abs(coef(summary(fittwo))[,5]), lower.tail = FALSE)

#final
fit1 = gee(Z ~ treat + Sex  + treat * Sex, family = binomial(link = "logit"),
           data=diarrhea, id=School, corstr = "exchangeable")
summary(fit1)
2* pnorm(abs(coef(summary(fit1))[,5]), lower.tail = FALSE)

## the estimate and se from above output
est = c(-1.664,-2.505, 0.191,-1.021)
se = c(0.0982,0.2317, 0.0985,0.3831)

## treatgirl
est_tg = est[2]+est[4]
se_tg = sqrt((se[2])^2 + 2*se[2]*se[4]*(1) +(se[4])^2)

# OR for treat in boy
ORb = exp(est[2])
lCIb = exp(est[2]-1.96*se[2])
uCIb = exp(est[2]+1.96*se[2])
b<- cbind(ORb,lCIb, uCIb)
# OR for treat in girl
ORg = exp(est_tg)
lCIg = exp(est_tg-1.96*se_tg)
uCIg = exp(est_tg+1.96*se_tg)
g<- cbind(ORg,lCIg, uCIg)
#OR table
resu<-rbind(b,g) 
resu
rownames(resu) = c("treatboy","treatgirl")
colnames(resu) = c("est","lower","upper")
round(resu,3)
```


\
__Question 2: Chlamydia Incidence__\

```{r}
cdc <- get(load('/Users/mengjiawei/Desktop/2020fall/525/Final/CDC.RData'))
dim(cdc)
### descriptive statistics
library(table1)
label(cdc$PBlack2000)<-"Percent black population from Census 2000"
label(cdc$PHisp2000)<-"Percent Hispanic population from Census 2000"
label(cdc$Cases)<-"Reported cases of chlamydia"
label(cdc$HHIncome2000)<-"Median household income (in $1,000) from Census 2000"
label(cdc$Population)<-"At-risk population size"

library(tidyverse)
#check duplicate and missing
nrow(cdc)
cdc = cdc %>% 
  distinct() %>%
  filter(!is.na(Cases))
nrow(cdc)

#find number of county enrolled
length(unique(cdc$FIPS)) #131

#check year
unique(cdc$Year)

# descriptive statistics
# table the income, black population, hispanic population in county level
table1(~ HHIncome2000 + PBlack2000 + PHisp2000, data = cdc 
            %>% arrange(FIPS,Year) %>% filter(!duplicated(FIPS)))
cdcf1 = cdc %>%
  group_by(Year) %>%
  summarise(nid = n())
table1(~nid,data=cdcf1)

#check the number of visit for each county
cdcf2 = cdc %>% 
  group_by(FIPS) %>% 
  summarise(nvisit=n())
table1(~ nvisit , data=cdcf2)

#fit model
#center year
cdc$Year <- cdc$Year - min(cdc$Year)

#generate rate
cdc$Rate <- cdc$Cases/cdc$Population


library(lmerTest)
#fit model
fit.ran = lmer(Rate~(1|FIPS)  + Year + HHIncome2000 + PBlack2000 + PHisp2000,
                data=cdc)
summary(fit.ran)

fit.ran2 = lmer(Rate~(1|FIPS)  + Year + HHIncome2000 + PBlack2000,
                data=cdc)
summary(fit.ran2)
```


__Question 3: Power Calculations__\
a)Regression model$y_{ij} =\mu + \theta_i + \beta_1\times intervention_{ij}+\epsilon_{ij}$. Code the intervention term: 0 = control, 1 = intervention, for each subject in each group, there are 3 observations.\

b)
```{r}
## Define model parameters:
m = 250
n = 3
rho = 0.3
tau2=12^2*rho
sigma2=12^2-tau2
alpha = 0.05

## A sequence of effect sizes (slope) to calculate power for
effs = seq(0, 5, by = 0.1)

## Set up design matrix
X = cbind (rep (1,m*3*n), c(rep(1, m*n), rep(c(0,1,1), m),
                            rep(c(0,0,1), m)))

## Set up covariance matrix
R0 = matrix(tau2,n,n)
diag(R0) <- sigma2 + tau2
V <- kronecker(diag(1,m*3), R0)

##Covariance of beta_hat and standard errors
VCOV = solve (t(X)%*%solve(V)%*%X)
SE = sqrt (diag (VCOV))

## From Slide 10
power =  pnorm ( qnorm (alpha/2) + abs(effs)/SE[2])
plot (power~effs, type = "l", lwd = 3, xlab = "Effect Size (Slope)")
abline(h = 0.8, col = 2, lty= 3)
MDE = (qnorm(0.8)-qnorm(alpha/2))*SE[2]
MDE
```

The minimum detectable effect size between intervention and control with 80% power and a two-sided type I error rate of 0.05 is 1.4.\

c)
If the investigator interested in estimating how the duration of the intervention may impact its effect, we will create new variable ‘time’ to indicate the duration of intervention for that observation. Code the intervention term: 0 = control, 1 = intervention, for each subject, there are 3 observations. Code the time term: 0 = intervention duration is 0 year, 1= intervention duration is 1 year, 2 = intervention duration is 2 year, 3 = intervention duration is 3 year. 
Regression model$y_{ij} =\mu + \theta_i + \beta_1\times intervention_{ij}+ \beta_2 \times time_{ij} +\epsilon_{ij}$. \

```{r}
## Define model parameters:
m = 250
n = 3
rho = 0.3
tau2=12^2*rho
sigma2=12^2-tau2
alpha = 0.05

## A sequence of effect sizes (slope) to calculate power for
effs = seq(0, 5, by = 0.1)

## Set up design matrix
X = cbind (rep (1,m*3*n), c(rep(1, m*n), rep(c(0,1,1), m),
                            rep(c(0,0,1), m)), c(rep(c(1,2,3), m), rep(c(0,1,2), m),
                            rep(c(0,0,1), m)))

## Set up covariance matrix
R0 = matrix(tau2,n,n)
diag(R0) <- sigma2 + tau2
V <- kronecker(diag(1,m*3), R0)

## Covariance of beta_hat and standard errors
VCOV = solve (t(X)%*%solve(V)%*%X)
SE = sqrt (diag (VCOV))

## From Slide 10
power = pnorm ( qnorm (alpha/2) + abs(effs)/SE[2])
plot (power~effs, type = "l", lwd = 3, xlab = "Effect Size (Slope)")
abline(h = 0.8, col = 2, lty= 3)
MDE = (qnorm(0.8)-qnorm(alpha/2))*SE[2]
MDE
```

The minimum detectable effect size between treatment and placebo with 80% power and a two-sided type I error rate of 0.05 is 2.06.\
